Supplemental Information: Remotely sensing river greenhouse gas exchange velocity using the SWOT satellite

### 2022-06-08 docs/intial_drafts/supp_v11.Rmd
first round of review, setting up for rerun on HPC

## Test S1: Estimating how hydraulically wide SWOT rivers are
To quantify the prevalence of hydraulically-wide, SWOT-observable rivers, we used the dataset of field-measured river hydraulics outlined in Section 2.1. That dataset has over 500,000 discrete measurements of river width, velocity, area, and discharge that were made by the United States Geological Survey (USGS) to calibrate streamgauge rating curves. Here, we describe how this dataset was filtered down to `r HG_swot$n` rivers.

@@

**Small-eddy models (Equation 5 from Main Test)**

@@

**Reynolds extension models (Equation 6 from Main Text)**

@@

Prior distributions are defined by their hyperparameters. For BIKER, prior distributions are formalized as truncated normal distributions of the log-transformed terms such that $log(X) \sim  \mathcal{N}(\mu, \sigma^2)$ for $\lambda < log(X) < \gamma$, using prior hyperparameters mean ($\mu$), standard deviation ($\sigma$), and upper ($\gamma$) and lower bounds ($\lambda$) for any parameter *X*. It is important to again stress, as we do in the main text, that BIKER prior hyperparameters are described using only SWOT data to be completely globally implementable. This use of the data to describe the priors is analogous to the 'empirical Bayes method' [@hoffFirstCourseBayesian2009].

@@

This leaves the $k_{600_t}$ hyperparameters to be defined. $k_{600_{\mu}}$ is set by invoking the hydraulic geometry (HG) relationships developed in Section 3.2.3 of the main text and Table S2 using the data from @brinkerhoffReconcilingAtaStationAtManyStations2019. We replaced both depth and velocity terms from our gas exchange model (Equation 7 in the Main Text) with these HG models, resulting in Equation S19 where Q is the mass-conserved streamflow for the river reach.

@@

More specifically, we use the 166 hydraulically-wide measurements in our field-measured $k_{600}$ dataset (Section 2 in the Main Text) and push 166 different Monte Carlo simulations through Equation S20. Note that each Monte Carlo simulation is itself 10,000 runs in order to obtain a distribution of $k_{600}$ estimates. For each of the 166 distributions of estimated $l_{600}$, we extract the standard deviation as the uncertainty. We then take the average of those standard deviations to be a reasonable estimate of $\sigma_k_{600}$. This ultimately provided a value of 1.1 for the log-transformed $\sigma_k_{600}$ and should be used when running BIKER on real SWOT data.

$$log_{k_{600}} \sim normal \left[ log(\beta_1), log(\sigma_{\beta_1}) \right] + \frac{1}{4}normal\left[log(\bar{U}), log(\sigma_{mannings})\right] \frac{7}{16}log(gS) + \frac{9}{16}log(H)  \mathbf{(S20)}$$

>>>>

## Test S1: Estimating how hydraulically wide SWOT rivers are
To quantify the prevalence of hydraulically-wide, SWOT-observable rivers, we used the dataset of field-measured river hydraulics outlined in Section 2.1. That dataset has over 500,000 discrete measurements of river width, velocity, area, and discharge that were made by the United States Geological Survey (USGS) to calibrate streamgauge rating curves. Here, we describe how this dataset was filtered down to `r HG_swot$n` SWOT-observable rivers.

@@

**Small-eddy models (Equation 3 from Main Test)**

@@

**Reynolds extension models (Equation 4 from Main Text)**

@@

Prior distributions are defined by their hyperparameters. For BIKER, prior distributions are formalized as truncated normal distributions of the log-transformed terms such that $log(X) \sim  \mathcal{N}(\mu, \sigma^2)$ for $\lambda < log(X) < \gamma$, using prior hyperparameters mean ($\mu$), standard deviation ($\sigma$), and upper ($\gamma$) and lower bounds ($\lambda$) for any parameter *X*. It is important to again stress, as we do in the Main Text, that BIKER prior hyperparameters are described using only SWOT data to be completely globally implementable. This use of the data to describe the priors is analogous to the 'empirical Bayes method' [@hoffFirstCourseBayesian2009].

@@

This leaves the $k_{600_t}$ hyperparameters to be defined. $k_{600_{\mu}}$ is set by invoking the hydraulic geometry (HG) relationships described in Section 3.2.3 of the main text and Table S2 using the data from @brinkerhoffReconcilingAtaStationAtManyStations2019. We replaced both depth and velocity terms from our gas exchange model (Equation 7 in the Main Text) with these HG models, resulting in Equation S19 where Q is the mass-conserved streamflow for the river reach.

@@

More specifically, we use the 166 hydraulically-wide measurements in our field-measured $k_{600}$ dataset (Section 2 in the Main Text) and push 166 different Monte Carlo simulations through Equation S20. Note that each Monte Carlo simulation is itself 10,000 runs in order to obtain a distribution of $k_{600}$ estimates. For each of the 166 distributions of estimated $k_{600}$, we extract the standard deviation as the uncertainty. We then take the average of those standard deviations to be a reasonable estimate of $\sigma_k_{600}$. This ultimately provided a value of 1.1 for the log-transformed $\sigma_k_{600}$ and should be used when running BIKER on real SWOT data.

$$log_{k_{600}} \sim \mathcal{N} \left[ log(\beta_1), log(\sigma_{\beta_1}) \right] + \frac{1}{4}\mathcal{N}\left[log(\bar{U}), log(\sigma_{mannings})\right] \frac{7}{16}log(gS) + \frac{9}{16}log(H)  \mathbf{(S20)}$$

### 2022-06-08 docs/intial_drafts/supp_v5.Rmd
first round of review, setting up for rerun on HPC

---
title: "Supplemental Information: Exploiting hydraulic geometry to remotely sense gas exchange velocity via the SWOT satellite"
author:
  - Craig B Brinkerhoff:
      email: cbrinkerhoff@umass.edu
      institute: UMass
      correspondence: true
  - Peter A Raymond:
      institute: Yale
  - Colin J Gleason:
      institute: UMass
  - Christopher J Zappa:
      institute: Lamont
  - Merritt H Harlan:
      institute: UMass
institute:
  - UMass: Department of Civil & Environmental Engineering, University of Massachusetts, Amherst, MA
  - Yale: School of the Environment, Yale University, New Haven, CT
  - Lamont: Lamont-Doherty Earth Observatory, Columbia University, Palisades, NY
bibliography: References.bib
csl: agu_citation_style.csl
output:
  word_document:
      pandoc_args:
        - '--lua-filter=scholarly-metadata.lua'
        - '--lua-filter=author-info-blocks.lua'
      reference_docx: word-styles-reference.docx

@@

>>>>

---
title: "Supplemental Information: Gas exchange in large rivers influenced by multiple turbulence scales: implications for remotely sensing gas exchange via the SWOT satellite"
author:
  - Craig B Brinkerhoff:
      email: cbrinkerhoff@umass.edu
      institute: UMass
      correspondence: true
  - Peter A Raymond:
      institute: Yale
  - Colin J Gleason:
      institute: UMass
  - Christopher J Zappa:
      institute: Lamont
  - Merritt H Harlan:
      institute: UMass
institute:
  - UMass: Department of Civil & Environmental Engineering, University of Massachusetts, Amherst, MA
  - Yale: School of the Environment, Yale University, New Haven, CT
  - Lamont: Lamont-Doherty Earth Observatory, Columbia University, Palisades, NY
bibliography: References.bib
csl: agu_citation_style.csl
output:
  word_document:
      pandoc_args:
        - '--lua-filter=scholarly-metadata.lua'
        - '--lua-filter=author-info-blocks.lua'
      reference_docx: word-styles-reference.docx

@@

## Figure S2
![Figure S2 Conceptual river reach model used in this study following McFLI logic [@gleasonTrackingRiverFlows2017] DOOOOOOOOO](cache/FCO2/Beaulieu_timeseries.jpg)

### 2022-06-08 docs/intial_drafts/supp_v9.Rmd
first round of review, setting up for rerun on HPC

**Small-eddy models**

$\epsilon = \epsilon_S$

@@

$\epsilon = \epsilon_D$

@@

**Reynolds extension models**

$\epsilon = \epsilon_S$

@@

Finally, recall that $\sigma_{k_{600}}$ refers to the total uncertainty inherent in equation 9, i.e. stemming from the Reynolds-extension model (equation 7) and Manning's equation for $\bar{U}$. For the purposes of this study, we are validating BIKER against equation 7 in the main text and so all uncertainties associated with the $\beta_1$ parameter are ignored and we only need to reflect the Manning's uncertainty in our specification of $\sigma_{k_{600}}^2$. Therefore, we take @hagemannBAMBayesianAMHGManning2017's estimated uncertainty from Manning's equation to infer streamflow from SWOT observations (0.25) and inflate it slightly to also account for the hydraulically-wide channel assumption and arrive at 0.30. This is the $\sigma_{k_{600}}$ used in this study.

In the scenario that BIKER is run on real SWOT data, $\sigma_k_{600}$ must reflect the full uncertainty implicit in equation 9 in the main text. This means we must also account for uncertainty from the $\beta_1$ parameter. Assuming perfect, no-error measurements are made by the SWOT satellite, the $k_{600}$ relative uncertainty is expressed for some set of hydraulic observations as equation S20. We use the 166 hydraulically-wide measurements in our field-measured $k_{600}$ dataset and run 166 different Monte Carlo simulations (each 10,000 runs) to obtain 166 different $\sigma_k_{600}$ terms. We then take the average of those to be a reasonable estimate of $\sigma_k_{600}$. This ultimately provided a value of 1.1 for the log-transformed $\sigma_k_{600}$ that should be used once SWOT launches.

@@

## Text S4: Literature $CO_2$ flux models
Here, we describe in more detail the $CO_2$ flux models used to compare against BIKER in section 3.2.2. The specific equations used are outlined in Table S3.

@@

For example, the 'Raymond 2012' model uses the hydraulic scaling equations from @raymondScalingGasTransfer2012 to predict velocity and depth, which are in turn used to calculate a $k_{CO_2}$, which finally is used to calculate $FCO_2$. Similarly, the 'Raymond 2013' model uses the equations outlined in @raymondGlobalCarbonDioxide2013 and the 'Brinkerhoff 2019' model uses new equations fit to the hydraulics dataset from @brinkerhoffReconcilingAtaStationAtManyStations2019 following Text S1. We chose to include the 'Brinkerhoff 2019' model as the training dataset is far larger than those used in either of the previous two models (Table S3: 104,624 versus 10,837 versus 1,026, respectively). Finally, we converted from $k_600$ to $k_{CO_2}$ following @raymondScalingGasTransfer2012.

@@

|**Description**|**Acronym**|**Definition**|**Ideal Score**|**Possible Range**|
|---------------|-----------|--------------|---------------|------------------|
|Correlation Coefficient|$r$| $\sqrt{1-(\frac{\sum\limits_{i=1}^{N_t}{(k_{600, i}-\hat{k_{600_i}})^2}}{\sum{(k_{600, i}-\bar{k_{600}})^2}})}$ |1|-1 to 1|
|Normalized root-mean-square error|NRMSE| $\sqrt{\frac{1}{N_t}\sum\limits_{i=1}^{N_t}(\frac{\hat{k_{600, i}}-k_{600, i}}{\bar{k_{600, i}}})^2}$ |0|0 to ∞|
|Kling-Gupta Efficiency|KGE|$1 - \sqrt{ (r(k_{600_i}, \hat{k_{600_i}}) - 1)^2 + (\frac{\mu_{\hat{k_{600_i}}} }{\mu_{k_{600_i}}} - 1)^2 + (\frac{ \sigma _{\hat{k_{600_i}}} / \mu _{\hat{k_{600_i}}} }{ \sigma_{k_{600_i}} / \mu_{k_{600_i}}} - 1)^2 }$|1|-∞ to 1|
|Mean Absolute Error|MAE| $\frac{\sum\limits_{i=1}^{N_t}(|\hat{k_{600_i}}-k_{600_i})|)}{n})$ |0|0 to ∞|

>>>>

**Small-eddy models**
**Log-law-of-the-wall model for the turbulent dissipation rate**

@@

**Form-drag model for the turbulent dissipation rate**

@@

**Reynolds extension models**
**Log-law-of-the-wall model for the turbulent dissipation rate**

@@

## Text S4: Detemrining complete BIKER model uncertainty
Recall that $\sigma_{k_{600}}$ refers to the total uncertainty inherent in equation 9, i.e. stemming from the Reynolds-extension model (equation 7) and Manning's equation for $\bar{U}$. For the purposes of this study, we are validating BIKER against equation 7 in the main text and so all uncertainties associated with the $\beta_1$ parameter are ignored and we only need to reflect the Manning's uncertainty in our specification of $\sigma_{k_{600}}^2$. Therefore, we take @hagemannBAMBayesianAMHGManning2017's estimated uncertainty from Manning's equation to infer streamflow from SWOT observations (0.25) and inflate it slightly to also account for the hydraulically-wide channel assumption and arrive at 0.30. This is the $\sigma_{k_{600}}$ used in this study.

In the scenario that BIKER is run on real SWOT data, $\sigma_k_{600}$ must reflect the full uncertainty implicit in equation 9 in the main text. This means we must also account for uncertainty from the $\beta_1$ parameter. Assuming perfect, no-error measurements are made by the SWOT satellite, the full equation 7 uncertainty is expressed for some set of hydraulic observations as equation S20. To quantify total uncertainty in $k_{600}$, we then push 166 different Monte Carlo simulations through equation S20 and take the mean uncertainty as a reasonable value for $\sigma_{k_{600}}$ once SWOT launches.

To do this, we use the 166 hydraulically-wide measurements in our field-measured $k_{600}$ dataset (section 2 in the Main Text) and push 166 different Monte Carlo simulations through equation S20. NOte that each Monte Carlo simulation is itself 10,000 runs in order to obtain a distribution of $k_{600}$ estimates. For each of the 166 distributions of estimated $l_{600}$, we extract the standard deviation. We then take the average of those standard deviations to be a reasonable estimate of $\sigma_k_{600}$. This ultimately provided a value of 1.1 for the log-transformed $\sigma_k_{600}$ that should be used once SWOT launches.

@@

## Text S5: Literature $CO_2$ flux models
Here, we describe in more detail the $CO_2$ flux models used to compare against BIKER in section 3.2.2. The specific equations used are outlined in Table S3.

@@

For example, the 'Raymond 2012' model uses the hydraulic scaling equations from @raymondScalingGasTransfer2012 to predict velocity and depth, which are in turn used to calculate a $k_{CO_2}$, which finally is used to calculate $FCO_2$. Similarly, the 'Raymond 2013' model uses the equations outlined in @raymondGlobalCarbonDioxide2013 and the 'Brinkerhoff 2019' model uses new equations fit to the hydraulics dataset from @brinkerhoffReconcilingAtaStationAtManyStations2019 after the filtering descirbed in Text S1. We chose to include the 'Brinkerhoff 2019' model as the training dataset is far larger than those used in either of the previous two models (Table S3: 104,624 versus 10,837 versus 1,026, respectively). Finally, we converted from $k_600$ to $k_{CO_2}$ following @raymondScalingGasTransfer2012.

@@

|**Description**|**Acronym**|**Definition**|**Ideal Score**|**Possible Range**|
|---------------|-----------|--------------|---------------|------------------|
|Correlation Coefficient|$r$| $\sqrt{1-(\frac{\sum\limits_{i=1}^{N_t}{(k_{600, i}-\hat{k_{600_i}})^2}}{\sum{(k_{600, i}-\mu_{k_{600}})^2}})}$ |1|-1 to 1|
|Normalized root-mean-square error|NRMSE| $\sqrt{\frac{1}{N_t}\sum\limits_{i=1}^{N_t}(\frac{\hat{k_{600, i}}-k_{600, i}}{\mu_{k_{600, i}}})^2}$ |0|0 to ∞|
|Kling-Gupta Efficiency|KGE|$1 - \sqrt{ (r(k_{600_i}, \hat{k_{600_i}}) - 1)^2 + (\frac{\mu_{\hat{k_{600}}} }{\mu_{k_{600}}} - 1)^2 + (\frac{ \sigma _{\hat{k_{600}}} / \mu _{\hat{k_{600}}} }{ \sigma_{k_{600}} / \mu_{k_{600}}} - 1)^2 }$|1|-∞ to 1|
|Normalized Mean Absolute Error|MAE| $\frac{\frac{\sum\limits_{i=1}^{N_t}(|\hat{k_{600_i}}-k_{600_i})|)}{n})}{\mu_{k_{600}}}$ |0|0 to ∞|

### 2022-06-08 manuscript_r_v1.Rmd
first round of review, setting up for rerun on HPC

---
title: "Remotely sensing river greenhouse gas exchange velocity using the SWOT satellite"
author:
  - Craig B Brinkerhoff:
      email: cbrinkerhoff@umass.edu
      institute: UMass
      correspondence: true
  - Colin J Gleason:
      institute: UMass
  - Christopher J Zappa:
      institute: Lamont
  - Peter A Raymond:
      institute: Yale
  - Merritt E Harlan:
      institute: UMass
institute:
  - UMass: Department of Civil & Environmental Engineering, University of Massachusetts, Amherst, MA
  - Lamont: Lamont-Doherty Earth Observatory, Columbia University, Palisades, NY
  - Yale: School of the Environment, Yale University, New Haven, CT
bibliography: References.bib
csl: agu_citation_style.csl
output:
  word_document:
    #  pandoc_args:
    #    - '--lua-filter=scholarly-metadata.lua'
    #    - '--lua-filter=author-info-blocks.lua'
      reference_docx: word-styles-reference.docx
---

>>>>

---
title: "Remotely sensing river greenhouse gas exchange velocity using the SWOT satellite"
author:
  - Craig B Brinkerhoff:
      email: cbrinkerhoff@umass.edu
      institute: UMass
      correspondence: true
  - Colin J Gleason:
      institute: UMass
  - Christopher J Zappa:
      institute: Lamont
  - Peter A Raymond:
      institute: Yale
  - Merritt E Harlan:
      institute: UMass
institute:
  - UMass: Department of Civil & Environmental Engineering, University of Massachusetts, Amherst, MA
  - Lamont: Lamont-Doherty Earth Observatory, Columbia University, Palisades, NY
  - Yale: School of the Environment, Yale University, New Haven, CT
bibliography: References.bib
csl: agu_citation_style.csl
output:redoc::redoc
  #word_document:
    #  pandoc_args:
    #    - '--lua-filter=scholarly-metadata.lua'
    #    - '--lua-filter=author-info-blocks.lua'
  #    reference_docx: word-styles-reference.docx
---

### 2022-06-08 supp_r_v1.Rmd
first round of review, setting up for rerun on HPC

---
title: "Supplemental Information: Remotely sensing river greenhouse gas exchange velocity using the SWOT satellite"
author:
  - Craig B Brinkerhoff:
      email: cbrinkerhoff@umass.edu
      institute: UMass
      correspondence: true
  - Colin J Gleason:
      institute: UMass
  - Christopher J Zappa:
      institute: Lamont
  - Peter A Raymond:
      institute: Yale
  - Merritt E Harlan:
      institute: UMass
institute:
  - UMass: Department of Civil & Environmental Engineering, University of Massachusetts, Amherst, MA
  - Lamont: Lamont-Doherty Earth Observatory, Columbia University, Palisades, NY
  - Yale: School of the Environment, Yale University, New Haven, CT
bibliography: References.bib
csl: agu_citation_style.csl
output:
  word_document:
  #    pandoc_args:
  #      - '--lua-filter=scholarly-metadata.lua'
  #      - '--lua-filter=author-info-blocks.lua'
      reference_docx: word-styles-reference.docx

>>>>

---
title: "Supplemental Information: Remotely sensing river greenhouse gas exchange velocity using the SWOT satellite"
author:
  - Craig B Brinkerhoff:
      email: cbrinkerhoff@umass.edu
      institute: UMass
      correspondence: true
  - Colin J Gleason:
      institute: UMass
  - Christopher J Zappa:
      institute: Lamont
  - Peter A Raymond:
      institute: Yale
  - Merritt E Harlan:
      institute: UMass
institute:
  - UMass: Department of Civil & Environmental Engineering, University of Massachusetts, Amherst, MA
  - Lamont: Lamont-Doherty Earth Observatory, Columbia University, Palisades, NY
  - Yale: School of the Environment, Yale University, New Haven, CT
bibliography: References.bib
csl: agu_citation_style.csl
output:redoc::redoc
  #word_document:
  #    pandoc_args:
  #      - '--lua-filter=scholarly-metadata.lua'
  #      - '--lua-filter=author-info-blocks.lua'
  #    reference_docx: word-styles-reference.docx
